An Efficient Interpolation Technique for Jump Proposals in Reversible-Jump Markov Chain Monte Carlo Calculations

### 2011-03-16 paper.tex
Minor edits to IV.

In this section, we demonstrate the efficiency of the kD-interpolated
jump proposal on a toy model comparison problem.  We draw $N = 100$
simulated data points from a $N(0,1)$ Gaussian distribution, and then
ask whether these data are better described by a model where they are
Gaussian distributed with unkonwn mean, $\mu$, and standard deviation,
$\sigma$
%
\be
p(x) = \frac{1}{\sqrt{2\pi} \sigma} \exp\left( - \frac{(x-\mu)^2}{2
    \sigma^2} \right),
\ee
%
or by a model where they are Cauchy distributed with mode $\alpha$,
and width $\beta$
%
\be
p(x) = \frac{1}{\pi \beta \left( 1 + \left(\frac{x - \alpha}{\beta}\right)^2\right)}.
\ee
%
We take priors on $\mu$ and $\alpha$ to be uniform in $[-1,1]$, and
priors in $\sigma$ and $\beta$ to be uniform in $[0.5, 1.5]$.  With a
data set of 100 points, the relative uncertainty in determining the
parameters of the underlying distribution is approximately 10\%, so we
expect the posterior probabilities in the $(\mu,\sigma)$ and
$(\alpha,\beta)$ spaces to occupy only a few percent of the prior
volume.  The Cauchy distribution is so much broader than the Gaussian
(it has no finite moments), that with uniform model proirs, the
posterior probability for the Gaussian model over the Cauchy model is
extremely large:
%
\be
\frac{p(\textnormal{Gaussian} | d)}{p(\textnormal{Cauchy}|d)} > 10^9
\ee
%

>>>>

In this section, we demonstrate the efficiency of the kD-interpolated
jump proposal on a toy model-comparison problem.  We draw $N = 100$
simulated data points from a $N(0,1)$ Gaussian distribution, and then
ask whether these data are better described by a model where they are
Gaussian distributed with unknown mean $\mu$ and standard deviation
$\sigma$
%
\be
p(x) = \frac{1}{\sqrt{2\pi} \sigma} \exp\left( - \frac{(x-\mu)^2}{2
    \sigma^2} \right)\ ,
\ee
%
or by a model where they are Cauchy distributed with mode $\alpha$
and width $\beta$
%
\be
p(x) = \frac{1}{\pi \beta \left( 1 + \left(\frac{x - \alpha}{\beta}\right)^2\right)}\ .
\ee
%
We take priors on $\mu$ and $\alpha$ to be uniform in $[-1,1]$, and
priors in $\sigma$ and $\beta$ to be uniform in $[0.5, 1.5]$.  With a
data set of 100 points, the relative uncertainty in determining the
parameters of the underlying distribution is approximately 10\%, so we
expect the posterior probabilities in the $(\mu,\sigma)$ and
$(\alpha,\beta)$ spaces to occupy only a few percent of the prior
volume.  The Cauchy distribution is much broader than the Gaussian
(it has no finite moments), so with uniform model priors, the
posterior probability for the Gaussian model over the Cauchy model is
extremely large:
%
\be
\frac{p(\textnormal{Gaussian} | d)}{p(\textnormal{Cauchy}|d)} > 10^9\ .
\ee
%
